Identifying Shifts in the Share Distribution of Transboundary Fish Stocks to Lead Time for Policy Response Under Climate Change

### 2020-12-22 Scripts/Manuscript.Rmd
las changes

The concept of shared stocks was developed following the ratification of the United Nations Convention on the Law of the Sea (UNCLOS) and the claiming of Exclusive Economic Zones (EEZs) by States [@UnitedNations:1986tl]. As defined by the United Nations’ Food and Agriculture Organization (FAO), shared stocks can be classified into four non-exclusive categories: (i) transboundary stocks, those that cross neighboring EEZs; (ii) straddling stocks, that, in addition to neighboring EEZs, also visit the adjacent high seas; (iii) highly migratory stocks, mainly tunas and bill-fishes, that migrate across vast oceanic regions including both the high seas and EEZs; and finally (iv) discrete stocks that are only present in the high seas [@Munro:2004th]. This study focuses on transboundary stocks exploited by fisheries operating within EEZs. While countries are responsible for the management of stocks within their EEZs, under UNCLOS, States are encouraged to cooperate when stocks are shared [@UnitedNations:1986tl]. A recent study estimates that, globally, there are 633 transboundary fish species representing 67% or the identified fished taxa, yielding an annual average of 48.5 million tonnes of catch and USD 78 billion in fishing revenue between 2005 and 2010 (Chapter 2).

The effectiveness of fisheries management for transboundary stocks is challenged by the shifts in marine stocks distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by]. In many cases, catch or fishing effort quotas for transboundary stocks are based on historical records [@Baudron:2020dj] and do not necessarily consider the biogeography of the stocks [@FredstonHermann:2018kp], nor the effects of climate change on the fish stocks and fisheries [@PalaciosAbrantes:2020tv; @Sumby:2021gw]. Misalignment between fisheries resources allocation and stocks’ distributional shifts have previously resulted in unsustainable harvest and international disputes [@Miller:2013iv; @Song:2017va; @Spijkers:2017ij], patterns that are expected to be exacerbated by intensifying climate change [@Pinsky:2018cb]. With future shifts in species distributions under climate change, the challenges for the fisheries management of transboundary stocks will continue to increase during the 21st century. In order to prepare ocean governance to meet the resource management challenges arising from shifting transboundary stocks, it is important to constrain when climate change will affect the dynamics of transboundary  stocks and the intensity of the resulting impacts [@Pinsky:2018cb; @PalaciosAbrantes:2020tv; @Link:2010ei].

The timing and intensity of climate impacts on managing transboundary fish stocks can be quantitatively examined through the concepts of ‘time of emergence’ (ToE) and ‘threat point’ (TP). Time of emergence is defined as the moment in time when a signal (e.g., future anthropogenic trend) emerges from the background noise of natural variability (e.g., historical natural variation) [@Hawkins:2012hc] and has been applied to multiple oceanic physical and biogeochemical variables [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Cheung:2020ja; @Schlunegger:2020gz; @Schlunegger:2019el; @Henson:2017ev]. The premise behind the time of emergence is that we can only be confident that a significant change has been detected when the signal of anthropogenic climate change is larger than the background noise of natural climate variability [@Hawkins:2012hc]. The concept of threat point comes from game theory and is defined as the minimum payoff a player is willing to receive to cooperate in a game theoretic model [@Nash:1953vr]. Game theory has been widely used to investigate and manage transboundary stocks [@Sumaila:2013hv], including the impacts of climate change on the economics of transboundary stocks between Canada and the United States (Sumaila et al., *in press*) .

Here, I estimate the time of emergence and threat point of the share distribution of transboundary stocks worldwide. I employ a species distribution model driven by output from a comprehensive Earth System Model with ten ensemble members to project the distribution of 663 transboundary species (Chapter 2) across 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario RCP 8.5 (see Methods). I treated each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population (Chapter 2) and only considered shared stocks between neighboring EEZs (i.e., transboundary stocks that exclude  the high seas. I find that many transboundary stocks have already ‘emerged’ from historical natural variability and exceeded their share threat point in most neighboring EEZs, posing a threat to their sustainability and the resilience of international treaties.

@@

This analysis includes 633 exploited marine transboundary species previously identified to account for 80% of the catch taken from the world’s EEZs between 2005 and 2014 (Chapter 2). Each time a species was shared by a pair of neighbouring EEZs it was considered a transboundary stock [@Teh:2015gd, Chapter 2], resulting in a total of 9,132 transboundary stocks in this analysis. Using the transboundary species Atlantic cod (*Gadus morhua*) as an example, and based on the definition of transboundary stock in this study, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share another stock of Atlantic cod, but the United States and Greenland do not share a stock of this transboundary species. I defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (Figure \@ref(fig:toeFigureS1) including island territories. I determined the intersections between polygons using the R-statistical software package sf [@PackagesfSimpleF:2018vp]. Each EEZ was categorized by geopolitical region according to the United Nations (https://population.un.org/wpp/DefinitionOfRegions/) and biome [@Reygondeau:2019uh]. The habitat association of each species was determined following the classification in FishBase for fish species (http://www.fishbase.org) and SeaLifeBase for invertebrates (http://www.fishbase.org). For each stock and EEZ, I used the Sea Around Us data to estimate catch and fishing revenue from fishing activities within global EEZs. I report both average catch and revenue for the last decade (2005-2014) [@Zeller2016; @Tai2017; @Sumaila:2015uc]. The monthly average US Consumer Price Index (CPI) according to the U.S. Bureau of Labor Statistic (https://www.bls.gov/cpi/) was used to standardize all monetary values to 2019 real USD.

@@

where $lat_{per}$ and $lon_{per}$ are the latitudes and longitudes of the grid cells holding the $per^{th}$ percentile of the projected transboundary stock abundance. To focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place, I included grid cells where the projected stock abundance within the neighbouring EEZs sharing the stock was above the top 95th percentile (e.g., per = 95%). A sensitivity analysis was ran using a subset of species (n = 34) for all EEZs (per = 20th, 50th and $90^{th}$ percentile) and examined the effects of different thresholds on the calculated index value (Figure \@ref(fig:toeFigureS3)). The centroid of each EEZ was estimated using the st package in R (Figure \@ref(fig:toeFigureS1)). For each ensemble member, neighboring EEZs and transboundary stock, I computed the distance between centroids assuming the earth is a perfect sphere and ignoring geographic barriers using the geosphere package in R;

@@

In game theory, cooperation over a shared resource will more likely result in a better overall outcome for the sharing parties than those operating on a non-cooperative basis [@Munro:1979dg]. However, the benefits that each player gets from a cooperative strategy must be above a minimum payoff i.e., “threat point” [@Sumaila:2013hv]. Here, I define the threat point as the minimum required proportion of a shared stock present within an EEZ for a country to engage in cooperative management with their sharing neighbor [@PalaciosAbrantes:2020vt; @Sumaila:2020tp]. Any proportion below the defined threat point of a country would result in a unilateral management of the shared stock.

First, I estimated the stock share ratio (SSR) of each transboundary stock shared by neighbouring nations between 1951 and 2100. I did this by aggregating the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock held within each EEZ (Chapter 2). Second, I averaged the calculated proportion of stock within an EEZ held into three time periods to reduce the effects of variability. The first time period ($SSR_{th}$) spans 1951 to 2005 and is chosen to match the historical period in the GFDL ESM2M simulations, over which the model was forced with observation-based greenhouse gas, aerosol and natural external forcing [@Dunne:2012bo; @Dunne:2013db]. It is assumed to represent the historic baseline of shared distributions for each stock. I then selected two future periods; the early $21^{st}$ century as the average of 2021 to 2040 ($SSR_{te}$), and the mid $21^{st}$ century as the average of 2041 to 2060 ($SSR_{tm}$). I chose these time periods to match the challenges of achieving fisheries-related UN-SDGs (6) such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) or SDG 1.2 (poverty reduction), to be achieved by 2030 [@Singh:2017ds]. The analysis was replicated for projected stocks distributions from each of the ten ensemble members and results were averaged across ensemble members. Third, I defined a threat point for each EEZ’s stock as $SSR_{th} \pm \sigma$, where $\sigma$ is the standard deviation of $SSR_{th}$. This way, a change of SSR beyond an EEZ’s threat point happens when the future SSR exceeds one standard deviation of the historical variations of the SSR i.e when $SSR_{te} \geqq (SSR_{th} + \sigma)$ or $SSR[t]_{e,m} \leqq (SSR_{th} - \sigma)$. Finally, I estimated the percentage change in the SSR ($\Delta{SSR}$) of each future time period ($SSR_f$) relative to the historic time period ($SSR_{th}$) for each stock whose share ratio overpassed the threat point following Palacios-Abrantes et al. (*in press*),

@@

```{r Load_Results, eval = TRUE,  echo = FALSE, include = FALSE}

@@

The median time of emergence of the transboundary index of stocks varies significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, DF = 93, p < 0.001; Figure \@ref(fig:toeFigure2)A). Overall, most tropical EEZs will see an earlier time of emergence with the EEZs of Latin America and the Caribbean and Polynesia having significantly earlier times of emergence (p < 0.05; see Table \@ref(tab:toeTableS1) for test statistics) than almost all other regions e.g., Ecuador (ToE = 2013) and Papa New Guinea (ToE = 2018) (Figure \@ref(fig:toeFigure2)B). In contrast, EEZs located in temperate regions like northern Europe and eastern Asia have significantly later times of emergence (p < 0.05; see Table \@ref(tab:toeTableS1) for test statistics) than the rest of the world, e.g., Ireland (ToE = 2067) and North Korea (ToE = 2058). However, notable exceptions to these broad patterns exist, with  tropical EEZs displaying a later average time of emergence across stocks including the  Pacific EEZ of Honduras (ToE = 2071) and Panama (ToE = 2070) and despite being outside the tropics, Belgium (ToE = 2015) and Norway (ToE = 2019) display a relatively early average time of emergence (Figure \@ref(fig:toeFigure2)A). There are only a few cases where the transboundary index of a stock does not emerge between 2006 and 2100. In the Arctic EEZs of Canada and the US, the present study included three stocks: capelin (*Mallotus villosus*), saffron cod (*Eleginus gracilis*), and Pacific herring (*Clupea pallasii*). In the case of Brazil, I included four transboundary stocks, all shared with Uruguay (Chapter 2): Argentine menhaden (*Brevoortia pectinata*), Argentine shortfin squid (*Illex argentinus*), Argentine shrimp (*Pleoticus muelleri*), and pink shrimp (*Penaeus paulensis*). However, on all cases, I find that the transboundary index of these stocks did not emerge before 2100 relative to the historic (1951-2005) distribution (Figure \@ref(fig:toeFigure2)A).

```{r toeFigure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Average time of emergence of stocks’ transboundary index. A) Land polygon shows the participation of the emerging stocks in the country’s total fishing revenue from transboundary stocks. Exclusive Economic Zone polygon represents the average time of emergence of the transboundary index across all stocks. Aqua color represents EEZs with no time of emergence between 2006 and 2100. B) Time of emergence of the transboundary index according to the United Nations sub-regions. Whiskers represent 1.5* interquartile range. Box represents interquartile range as distance between first and third quartiles. Line represents median, and black points represent outliers with values exceeding of 1.5 times the interquartile range."}

@@

```{r toeFigure3, eval = T,  echo = F, fig.width=7, fig.height=8,fig.pos= "H", fig.cap =  "Average time of emergence of stocks’ transboundary index per fishing entity, aggregated by colour and shape according to region. Fishing revenue on a logarithmic scale. Showing country names for the top 90th percentile of annual fishing revenue."}

@@

I set an arbitrary threshold of 1 s.d. (67% confidence) to determine the time of emergence of the transboundary index and the change in stock share ratio below the threat point (see Methods). I tested the sensitivity to a larger threshold for the time of emergence (e.g., 2 s.d for a 95% confidence) and the global average time of emergence of the transboundary index considering a 2 s.d threshold was 2036 $\pm$ 28, a difference of about 7 years relative to the 1 s.d. threshold (Figure \@ref(fig:toeFigureS2)). In total, 4,119 transboundary stocks would emerge between 2006 and 2100 considering a more conservative metric,

I compared stocks time of emergence within EEZs by the species’ habitat association (Figure \@ref(fig:toeFigure4)). Each species is assigned to be associated to one of the seven habitat types based on information available from FishBase and SeaLifeBase for fish and invertebrates, respectively (see Methods). I found significant differences in the time of emergence of stocks range shifts, indicated by the transboundary index, by their habitat association (Kruskal-Wallis, $X^2$ = 286.48, DF = 93, p < 0.001). Specifically, the time of emergence of the transboundary index of species that are associated with pelagic-oceanic and reef habitats were significantly different than other habitat preferences (See Table \@ref(tab:toeTableS2)). On average, species associated with pelagic oceanic habitats e,g, tunas (*Thunnus sp.*), bathy-pelagic habitats e.g., blue whiting (*Micromesistius poutassou*) and corals (e.g., Bigeye snapper *Lutjanus lutjanus*) are projected to have an earliest time of emergence (Figure \@ref(fig:toeFigure4)). Species associated with bathydemersal habitats such as Alaska plaice (*Pleuronectes quadrituberculatus*) generally have an average time of emergence of range shifts that is later than 2025 (Figure \@ref(fig:toeFigure4)).

```{r toeFigure4, eval = T,  echo = F, fig.width=12, fig.height=7,fig.pos= "H", fig.cap =  "Comparison of regional time of emergence of shared stocks’ range shifts by species’ habitat association. Classification is based on the habitat preference information obtained from FishBase and SeaLifeBase (see Methods). The number of species included in this analysis for each habitat type is noted in parenthesis. Whiskers represent 1.5* interquartile range. Box represents interquartile range as distance between first and third quartiles. Line represents median, and black points represent outliers with values exceeding 1.5 times the interquartile range."}

@@

```{r toeFigure5, eval = T,  echo = F, fig.width=12, fig.height=10,fig.pos= "H", fig.cap =  "Changes in stock share ratio below each country’s threat point by 2030 (2021-2040) relative to 1951-2005. Lines represent the average gain of transboundary stock share ratio with arrows going from EEZ with decreasing stock share (point) to those that are gaining shares (arrowhead). Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point. Panel B zooms in to specific areas shown within the grey dashed line boundaries in A."}

@@

Our projections for the emergence of distribution shifts of transboundary stocks in the early $21^{st}$ century concur with previous studies that have detected and attributed changes in marine catch composition [@Cheung:2013gk]. For example, in the early 2000s, Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range poleward, reaching the coast of Washington state (US). A new fishery targeting Humboldt squid was developed quickly following its range expansion [@Pinsky:2014by; @Zeidberg:2007gs]. In the northeast Atlantic, the Atlantic mackerel (*Scomber scombrus*) fisheries are multi-laterally managed by the EU, Norway, Iceland, Russia, and Denmark (on behalf of the Faroe Islands and Greenland) through the North-East Atlantic Fisheries Commission (NEAFC). However, a range expansion of Atlantic mackerel into Icelandic waters in 2007 resulted in Iceland capturing 6% of the total fishery catch and a further 18% capture in 2008, without consultation with NEAFC and threatening the sustainability of the stock [@Spijkers:2017ij]. Such changes resulted in disputes between Iceland and the Faroes Islands, as well as the NEAFC member states [@Spijkers:2017ij]. As climate change impacts on transboundary stocks continue to emerge from historical variability in the coming years, mal-adapted fisheries management and international disputes like that of Atlantic mackerel are likely to increase in frequency [@Pinsky:2018cb].

@@

Our results underscore regional ‘hotspots’ of climate risk for transboundary fisheries management that will require the prompt adaptation of management plans (Figure \@ref(fig:toeFigure2)). For example, Polynesia and Micronesia are the only regions where the average time of emergence of the distribution of where all transboundary stocks across all EEZs emerged before 2022 (Figure \@ref(fig:toeFigure5)). Moreover, countries in these hotspot regions are highly dependent on pelagic-oceanic (e.g., tunas and bill fishes; [@Johnson:2017ev]) and reef-associated transboundary stocks [@Cabral:2018hc] for both food security and profits [@CisnerosMontemayor:2016gq; @Hanich:2018dh]. On the other hand, “coldspots” of climate risk exist for transboundary fisheries, like the Northeast Atlantic and Northwest Pacific, where stocks are expected to emerge later in the $21^{st}$ century. Assuming that these countries already have adequate transboundary management to deal with historical and current variations in fish stocks’ distribution e.g., the Pacific Salmon Treaties [@Miller:2004fpa], management agencies are likely to have more time to prepare for emerging stock shifts.

@@

```{r FigureS1, eval = F,  echo = F, fig.width=9, fig.height=4, fig.cap =  "World Exclusive Economic Zones used in this study as devided by the Sea Around Us and their centroids (points)."}

>>>>

The concept of shared stocks was developed following the ratification of the United Nations Convention on the Law of the Sea (UNCLOS) and the claiming of Exclusive Economic Zones (EEZs) by States [@UnitedNations:1986tl]. As defined by the United Nations’ Food and Agriculture Organization (FAO), shared stocks can be classified into four non-exclusive categories: (i) transboundary stocks, those that cross neighboring EEZs; (ii) straddling stocks, that, in addition to neighboring EEZs, also visit the adjacent high seas; (iii) highly migratory stocks, mainly tunas and bill-fishes, that migrate across vast oceanic regions including both the high seas and EEZs; and finally (iv) discrete stocks that are only present in the high seas [@Munro:2004th]. This chapter focuses on transboundary stocks exploited by fisheries operating within EEZs. While countries are responsible for the management of stocks within their EEZs, under UNCLOS, States are encouraged to cooperate when stocks are shared [@UnitedNations:1986tl]. In the previous chapter, I estimated that globally, there are 633 transboundary fish species representing 67% or the identified fished taxa, yielding an annual average of 48.5 million tonnes of catch and USD 78 billion in fishing revenue between 2005 and 2010.

The effectiveness of fisheries management for transboundary stocks is challenged by the shifts in marine stocks distribution under climate change [@Pinsky:2018cb; @Pinsky:2014by]. In many cases, catch or fishing effort quotas for transboundary stocks are based on historical records [@Baudron:2020dj] and do not necessarily consider the biogeography of the stocks [@FredstonHermann:2018kp], nor the effects of climate change on the fish stocks and fisheries [@Sumby:2021gw, Chapter 2]. Misalignment between fisheries resources allocation and stocks’ distributional shifts have previously resulted in unsustainable harvest and international disputes [@Miller:2013iv; @Song:2017va; @Spijkers:2017ij], patterns that are expected to be exacerbated by intensifying climate change [@Pinsky:2018cb]. With future shifts in species distributions under climate change, the challenges for the fisheries management of transboundary stocks will continue to increase during the 21st century. In order to prepare ocean governance to meet the resource management challenges arising from shifting transboundary stocks, it is important to constrain when climate change will affect the dynamics of transboundary  stocks and the intensity of the resulting impacts [@Pinsky:2018cb; @Link:2010ei, Chapter 2].

The timing and intensity of climate impacts on managing transboundary fish stocks can be quantitatively examined through the concepts of ‘time of emergence’ (ToE) and ‘threat point’ (TP). Time of emergence is defined as the moment in time when a signal (e.g., future anthropogenic trend) emerges from the background noise of natural variability (e.g., historical natural variation) [@Hawkins:2012hc] and has been applied to multiple oceanic physical and biogeochemical variables [@Rodgers:2015jq; @Keller:2014jv; @Frolicher:2016dg; @Cheung:2020ja; @Schlunegger:2020gz; @Schlunegger:2019el; @Henson:2017ev]. The premise behind the time of emergence is that we can only be confident that a significant change has been detected when the signal of anthropogenic climate change is larger than the background noise of natural climate variability [@Hawkins:2012hc]. The concept of threat point comes from game theory and is defined as the minimum payoff a player is willing to receive to cooperate in a game theoretic model [@Nash:1953vr]. Game theory has been widely used to investigate and manage transboundary stocks [@Sumaila:2013hv], including the impacts of climate change on the economics of transboundary stocks between Canada and the United States [@Sumaila:2020tp].

Here, I estimate the time of emergence and threat point of the share distribution of transboundary stocks worldwide. I employ a species distribution model driven by output from a comprehensive Earth System Model with ten ensemble members to project the distribution of 663 transboundary species (Chapter 2) across 280 EEZs of 198 coastal countries under a high greenhouse gas emissions scenario RCP 8.5 (see Methods). I treated each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population (Chapter 2) and only considered shared stocks between neighboring EEZs (i.e., transboundary stocks that exclude the high seas). I find that the distribution of many transboundary stocks have already ‘emerged’ from historical natural variability and exceeded their share threat point in most neighboring EEZs, posing a threat to their sustainability and the resilience of international treaties.

@@

This analysis includes 633 exploited marine transboundary species previously identified to account for 80% of the catch taken from the world’s EEZs between 2005 and 2014 (Chapter 2). Each time a species was shared by a pair of neighbouring EEZs it was considered a transboundary stock [@Teh:2015gd, Chapter 2], resulting in a total of 9,132 transboundary stocks in this analysis. Using the transboundary species Atlantic cod (*Gadus morhua*) as an example, and based on the definition of transboundary stock in this chapter, the United States and Canada share a stock of Atlantic cod, and Canada and Greenland share another stock of Atlantic cod, but the United States and Greenland do not share a stock of this transboundary species. I defined the boundaries of the world’s EEZs using the Sea Around Us spatial division (updated 1 July 2015, available from http://www.seaaroundus.org), noting that it subdivides the EEZs of 198 coastal states into 280 regions (Figure \@ref(fig:toeFigureS1)) including island territories. I determined the intersections between polygons using the R statistical software package sf [@PackagesfSimpleF:2018vp]. Each EEZ was categorized by geopolitical region according to the United Nations (https://population.un.org/wpp/DefinitionOfRegions/) and biome [@Reygondeau:2019uh]. The habitat association of each species was determined following the classification in FishBase for fish species (http://www.fishbase.org) and SeaLifeBase for invertebrates (http://www.fishbase.org). For each stock and EEZ, I used the Sea Around Us data to estimate catch and fishing revenue from fishing activities within global EEZs. I report both average catch and revenue for the last decade (2005-2014) [@Zeller2016; @Tai2017; @Sumaila:2015uc]. The monthly average US Consumer Price Index (CPI) according to the U.S. Bureau of Labor Statistic (https://www.bls.gov/cpi/) was used to standardize all monetary values to 2019 real USD.

@@

where $lat_{per}$ and $lon_{per}$ are the latitudes and longitudes of the grid cells holding the $per^{th}$ percentile of the projected transboundary stock abundance. To focus on areas where transboundary stocks are more abundant and fishing activities are more likely to take place, I included grid cells where the projected stock abundance within the neighbouring EEZs sharing the stock was above the top 95th percentile (e.g., per = 95%). A sensitivity analysis was ran using a subset of species (n = 34) for all EEZs (per = $20^{th}$, $50^{th}$ and $90^{th}$ percentile) and examined the effects of different thresholds on the calculated index value (Figure \@ref(fig:toeFigureS3)). The centroid of each EEZ was estimated using the st package in R (Figure \@ref(fig:toeFigureS1)). For each ensemble member, neighboring EEZs and transboundary stock, I computed the distance between centroids assuming the earth is a perfect sphere and ignoring geographic barriers using the geosphere package in R;

@@

In game theory, cooperation over a shared resource will more likely result in a better overall outcome for the sharing parties than those operating on a non-cooperative basis [@Munro:1979dg]. However, the benefits that each player gets from a cooperative strategy must be above a minimum payoff i.e., “threat point” [@Sumaila:2013hv]. Here, I define the threat point as the minimum required proportion of a shared stock present within an EEZ for a country to engage in cooperative management with their sharing neighbor [Chapter 2, @Sumaila:2020tp]. Any proportion below the defined threat point of a country would result in a unilateral management of the shared stock.

First, I estimated the stock share ratio (SSR) of each transboundary stock shared by neighbouring nations between 1951 and 2100. I did this by aggregating the number of $0.5^{\circ}$ x $0.5^{\circ}$ grid cells in which the stock is present across neighboring EEZs, and then calculated the proportion of the stock held within each EEZ (Chapter 2). Second, I averaged the calculated proportion of stock within an EEZ held into three time periods to reduce the effects of variability. The first time period ($SSR_{th}$) spans 1951 to 2005 and is chosen to match the historical period in the GFDL ESM2M simulations, over which the model was forced with observation-based greenhouse gas, aerosol and natural external forcing [@Dunne:2012bo; @Dunne:2013db]. It is assumed to represent the historic baseline of shared distributions for each stock. I then selected two future periods; the early $21^{st}$ century as the average of 2021 to 2040 ($SSR_{te}$), and the mid $21^{st}$ century as the average of 2041 to 2060 ($SSR_{tm}$). I chose these time periods to match the challenges of achieving fisheries-related UN-SDGs such as SDG 14.4 (end overfishing), SDG 2.4 (ensure sustainable food production systems) or SDG 1.2 (poverty reduction), to be achieved by 2030 [@Singh:2017ds]. The analysis was replicated for projected stocks distributions from each of the ten ensemble members and results were averaged across ensemble members. Third, I defined a threat point for each EEZ’s stock as $SSR_{th} \pm \sigma$, where $\sigma$ is the standard deviation of $SSR_{th}$. This way, a change of SSR beyond an EEZ’s threat point happens when the future SSR exceeds one standard deviation of the historical variations of the SSR i.e when $SSR_{te} \geqq (SSR_{th} + \sigma)$ or $SSR[t]_{e,m} \leqq (SSR_{th} - \sigma)$. Finally, I estimated the percentage change in the SSR ($\Delta{SSR}$) of each future time period ($SSR_f$) relative to the historic time period ($SSR_{th}$) for each stock whose share ratio overpassed the threat point following (See Chapter 4 - Methods),

@@

```{r Load_Results, eval = F,  echo = FALSE, include = FALSE}

@@

The median time of emergence of the transboundary index of stocks varies significantly according to the geographic region of the neighboring EEZs (Kruskal-Wallis, $X^2$ = 287.23, DF = 93, p < 0.001; Figure \@ref(fig:toeFigure2)A). Overall, most tropical EEZs will see an earlier time of emergence with the EEZs of Latin America and the Caribbean and Polynesia having significantly earlier times of emergence (p < 0.05; see Table \@ref(tab:toeTableS1) for test statistics) than almost all other regions e.g., Ecuador (ToE = 2013) and Papa New Guinea (ToE = 2018) (Figure \@ref(fig:toeFigure2)B). In contrast, EEZs located in temperate regions like northern Europe and eastern Asia have significantly later times of emergence (p < 0.05; see Table \@ref(tab:toeTableS1) for test statistics) than the rest of the world, e.g., Ireland (ToE = 2067) and North Korea (ToE = 2058). However, notable exceptions to these broad patterns exist, with  tropical EEZs displaying a later average time of emergence across stocks including the  Pacific EEZ of Honduras (ToE = 2071) and Panama (ToE = 2070) and despite being outside the tropics, Belgium (ToE = 2015) and Norway (ToE = 2019) display a relatively early average time of emergence (Figure \@ref(fig:toeFigure2)A). There are only a few cases where the transboundary index of a stock does not emerge between 2006 and 2100. In the Arctic EEZs of Canada and the US, the present chapter included three stocks: capelin (*Mallotus villosus*), saffron cod (*Eleginus gracilis*), and Pacific herring (*Clupea pallasii*). In the case of Brazil, I included four transboundary stocks, all shared with Uruguay (Chapter 2): Argentine menhaden (*Brevoortia pectinata*), Argentine shortfin squid (*Illex argentinus*), Argentine shrimp (*Pleoticus muelleri*), and pink shrimp (*Penaeus paulensis*). However, on all cases, I find that the transboundary index of these stocks did not emerge before 2100 relative to the historic (1951-2005) distribution (Figure \@ref(fig:toeFigure2)A).

```{r toeFigure2, eval = T,  echo = F, fig.width  = 13,fig.pos= "H", fig.height =11, fig.cap =  "Average time of emergence of stocks’ transboundary index. A) Land polygon shows the participation of the emerging stocks in the country’s total fishing revenue from transboundary stocks. Exclusive Economic Zone polygon represents the average time of emergence of the transboundary index across all stocks. Aqua color represents EEZs with no time of emergence between 2006 and 2100. B) Time of emergence of the transboundary index according to the United Nations sub-regions. Whiskers represent 1.5* interquartile range. Box represents interquartile range as distance between first and third quartiles. Line represents median, and black points represent outliers with values exceeding of 1.5 times the interquartile range.", fig.scap="Average time of emergence of stocks’ transboundary index.", out.extra=''}

@@

```{r toeFigure3, eval = T,  echo = F, fig.width=7, fig.height=8,fig.pos= "H", fig.cap =  "Average time of emergence of stocks’ transboundary index per fishing entity, aggregated by colour and shape according to region. Fishing revenue on a logarithmic scale. Showing country names for the top 90th percentile of annual fishing revenue.", fig.scap="Average time of emergence of stocks’ transboundary index per fishing entity, aggregated by colour and shape according to region", out.extra=''}

@@

I set an arbitrary threshold of 1 s.d. (67% confidence) to determine the time of emergence of the transboundary index and the change in stock share ratio below the threat point (see Methods). I tested the sensitivity to a larger threshold for the time of emergence (e.g., 2 s.d for a 95% confidence) and the global average time of emergence of the transboundary index considering a 2 s.d threshold was 2036 $\pm$ 28, a difference of about 7 years relative to the 1 s.d. threshold (Figure \@ref(fig:toeFigureS2)). In total, 4,119 transboundary stocks would emerge between 2006 and 2100 considering a more conservative metric.

I compared stocks time of emergence within EEZs by the species’ habitat association (Figure \@ref(fig:toeFigure4)). Each species is assigned to be associated to one of the seven habitat types based on information available from FishBase and SeaLifeBase for fish and invertebrates, respectively (see Methods). I found significant statistical differences in the time of emergence of stocks range shifts, indicated by the transboundary index, by their habitat association (Kruskal-Wallis, $X^2$ = 286.48, DF = 93, p < 0.001). Specifically, the time of emergence of the transboundary index of species that are associated with pelagic-oceanic and reef habitats were significantly different than other habitat preferences (See Table \@ref(tab:toeTableS2)). On average, species associated with pelagic oceanic habitats e,g, tunas (*Thunnus sp.*), bathy-pelagic habitats e.g., blue whiting (*Micromesistius poutassou*) and corals (e.g., Bigeye snapper *Lutjanus lutjanus*) are projected to have an earliest time of emergence (Figure \@ref(fig:toeFigure4)). Species associated with bathydemersal habitats such as Alaska plaice (*Pleuronectes quadrituberculatus*) generally have an average time of emergence of range shifts that is later than 2025 (Figure \@ref(fig:toeFigure4)).

```{r toeFigure4, eval = T,  echo = F, fig.width=12, fig.height=6,fig.pos= "H", fig.cap =  "Comparison of regional time of emergence of shared stocks’ range shifts by species’ habitat association. Classification is based on the habitat preference information obtained from FishBase and SeaLifeBase (see Methods). The number of species included in this analysis for each habitat type is noted in parenthesis. Whiskers represent 1.5* interquartile range. Box represents interquartile range as distance between first and third quartiles. Line represents median, and black points represent outliers with values exceeding 1.5 times the interquartile range.", fig.scap="Comparison of regional time of emergence of shared stocks’ range shifts by species’ habitat association.", out.extra=''}

@@

```{r toeFigure5, eval = T,  echo = F, fig.width=12, fig.height=10,fig.pos= "H", fig.cap =  "Changes in stock share ratio below each country’s threat point by 2030 (2021-2040) relative to 1951-2005. Lines represent the average gain of transboundary stock share ratio with arrows going from EEZ with decreasing stock share (point) to those that are gaining shares (arrowhead). Land polygons represent the percentage of stocks that are projected to change their stock share ratio beyond the identified threat point. Panel B zooms in to specific areas shown within the grey dashed line boundaries in A.", fig.scap="Changes in stock share ratio below each country’s threat point by 2030 (2021-2040) relative to 1951-2005.", out.extra=''}

@@

My projections for the emergence of distribution shifts of transboundary stocks in the early $21^{st}$ century concur with previous studies that have detected and attributed changes in marine catch composition [@Cheung:2013gk]. For example, in the early 2000s, Humboldt squid (*Dosidicus gigas*) substantially expanded its geographic range poleward, reaching the coast of Washington state (US). A new fishery targeting Humboldt squid was developed quickly following its range expansion [@Pinsky:2014by; @Zeidberg:2007gs]. In the northeast Atlantic, the Atlantic mackerel (*Scomber scombrus*) fisheries are multi-laterally managed by the EU, Norway, Iceland, Russia, and Denmark (on behalf of the Faroe Islands and Greenland) through the North-East Atlantic Fisheries Commission (NEAFC). However, a range expansion of Atlantic mackerel into Icelandic waters in 2007 resulted in Iceland capturing 6% of the total fishery catch and a further 18% capture in 2008, without consultation with NEAFC and threatening the sustainability of the stock [@Spijkers:2017ij]. Such changes resulted in disputes between Iceland and the Faroes Islands, as well as the NEAFC member states [@Spijkers:2017ij]. As climate change impacts on transboundary stocks continue to emerge from historical variability in the coming years, mal-adapted fisheries management and international disputes like that of Atlantic mackerel are likely to increase in frequency [@Pinsky:2018cb].

@@

My results underscore regional ‘hotspots’ of climate risk for transboundary fisheries management that will require the prompt adaptation of management plans (Figure \@ref(fig:toeFigure2)). For example, Polynesia and Micronesia are the only regions where the average time of emergence of the distribution of where all transboundary stocks across all EEZs emerged before 2022 (Figure \@ref(fig:toeFigure5)). Moreover, countries in these hotspot regions are highly dependent on pelagic-oceanic (e.g., tunas and bill fishes; [@Johnson:2017ev]) and reef-associated transboundary stocks [@Cabral:2018hc] for both food security and profits [@CisnerosMontemayor:2016gq; @Hanich:2018dh]. On the other hand, “coldspots” of climate risk exist for transboundary fisheries, like the Northeast Atlantic and Northwest Pacific, where stocks are expected to emerge later in the $21^{st}$ century. Assuming that these countries already have adequate transboundary management to deal with historical and current variations in fish stocks’ distribution e.g., the Pacific Salmon Treaties [@Miller:2004fpa], management agencies are likely to have more time to prepare for emerging stock shifts.

@@

```{r FigureS1, eval = F,  echo = F, fig.width=9, fig.height=4, fig.cap =  "World Exclusive Economic Zones used in this chapter as devided by the Sea Around Us and their centroids (points)."}
